---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# PackageMLpredictions

<!-- badges: start -->
<!-- badges: end -->

`learnMET` (learn **M**ulti-**E**nvironment **T**rials) provides an integrated pipeline to: (1) facilitate environmental characterization with several options regarding the aggregation of daily weather data; and (2) evaluate various types of state-of-the-art machine learning approaches with respect to their predictive performance based on relevant cross-validation schemes for MET datasets (3) implementing predictions for unobserved configurations of genotypic and environmental predictors that the user
wants to test \textit{in silico}.

# Installation

Install the development version from [GitHub](https://github.com/cjubin/learnMET) with:

```{r, eval=FALSE}
devtools::install_github("cjubin/learnMET")

# To build the HTML vignette use
devtools::install_github("cjubin/learnMET", build_vignettes = TRUE)
```

